### 2019-08-14 Docs/Langdokumentation.tex
Added images and packages

\documentclass[11pt]{article}

% pdfLaTeX input and font settings
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage[english,ngerman]{babel}                 % German and English word separation settings
\usepackage[autostyle=true,german=quotes]{csquotes} % Better German quotes
\usepackage{graphicx}                               % Include graphics/images
\usepackage{subcaption}                             % Used for 2 images side-by-side
\usepackage[pdfa]{hyperref}                         % Klickable Links and ToC entries (should be last package !!!)

\title{DeepRain}
\author{Thomas Gnaedig, Etienne Gramlich, Tim, Merle Wolf}

\begin{document}
\maketitle
\tableofcontents

% Abstract

\input{Daten.tex}

% Datenaufbereitung

% Datenanalyse

% Netzwerke
\input{Architekturen.tex}
\input{Klassifizierung.tex}

% Fazit / Ausblick

\end{document}