### 2019-08-14 Docs/Langdokumentation.tex
Made Links black

\documentclass[11pt]{article}

% pdfLaTeX input and font settings
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage[english,ngerman]{babel}                 % German and English word separation settings
\usepackage[autostyle=true,german=quotes]{csquotes} % Better German quotes
\usepackage{graphicx}                               % Include graphics/images
\usepackage{subcaption}                             % Used for 2 images side-by-side
\usepackage[bookmarks,                 % Creates bookmarks
            bookmarksnumbered=true,    % Numbers in PDF bookmarks
            bookmarksopen=true,        % Don't show bookmaks when opened
            hyperindex=true,           % Numbers in index are links
            pdfpagelabels=true,        % Set PDF page labels
            pdfa]{hyperref}                     % Klickable Links and ToC entries (should be last package !!!)

\hypersetup{
    pdfauthor           = {Thomas Gnaedig, Etienne Gramlich, Tim, Merle Wolf},
    pdftitle            = {DeepRain},
    pdfsubject          = {Deep Learning},
    pdfkeywords         = {Deep Learning,rain},
    pdflang             = {de},         % Set language of document to German
    pdfstartview        = {FitH},       % Fit the document to screen horizontally (page width)
    pdfview             = {FitH},       % Fit the document to screen horizontally (page width)
    colorlinks          = {true},       % Use color for links
    linkcolor           = {black},      % Links in black (text color)
    urlcolor            = {black},      % URLs in black (text color)
    citecolor           = {black},      % Citations in black (text color)
    filecolor           = {black},      % File links in black (text color)
}

\title{DeepRain}
\author{Thomas Gnaedig, Etienne Gramlich, Tim, Merle Wolf}

\begin{document}
\maketitle
\tableofcontents

% Abstract

\input{Daten.tex}

% Datenaufbereitung

% Datenanalyse

% Netzwerke
\input{Architekturen.tex}
\input{Klassifizierung.tex}

% Fazit / Ausblick

\end{document}