### 2019-08-20 Docs/paper/report.tex

\documentclass[oneside]{htwg-report}

\include{header/header}

\addbibresource{./bib/report.bib}

\begin{document}

\pagenumbering{gobble}

%% 'reporttype' add background elements to the cover / front page
%% possible values are:
%% bachelor	--> B S C
%% master	--> M S C
%% other		--> none
\reporttype{master}

\reporttypetext{Teamprojekt (Master 3. Semester)}

\newcommand{\verfasserA}{Thomas Gnädig}
\newcommand{\verfasserB}{Etienne Gramlich}
\newcommand{\verfasserC}{Merle Wolf}
\newcommand{\verfasserD}{Tim}
\newcommand{\verfasserE}{}
\newcommand{\thema}{Teamprojektthema}
\newcommand{\hoschschule}{Hochschule für Technik, Wirtschaft und Gestaltung}
\newcommand{\institut}{HTWG Konstanz, Institut für Optische Systeme}
\newcommand{\prueferA}{Oliver Dürr}
\newcommand{\prueferB}{Betreuer B}

\title[Teamprojektthema]{\thema}

\doclocation{Konstanz}
\docdate{20. August 2018}

\makecover[]
%          
%% Include an optional title page.
%\input{title}

\input{extended_abstract}

\end{document}