### 2019-09-09 Docs/Langdokumentation.tex
Fazit

\documentclass[11pt,a4paper]{article}

% pdfLaTeX input and font settings
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}

\usepackage[english,ngerman]{babel}                 % German and English word separation settings
\newcommand{\gqq}[1]{\glqq #1\grqq{}}               % Own command for german quotes
\usepackage{graphicx}                               % Include graphics/images
\usepackage{float}
\usepackage{multirow}                               % Row spanning over multiple rows
\usepackage{subcaption}                             % Used for 2 images side-by-side
\usepackage{tabularx}                               % for 'tabularx' environment
\usepackage[font={small}]{caption}
\usepackage[printonlyused,withpage]{acronym}        % Abkürzungsverzeichnis, https://de.wikibooks.org/wiki/LaTeX-W%C3%B6rterbuch:_Abk%C3%BCrzungsverzeichnis
\usepackage[bookmarks,                 % Creates bookmarks
            bookmarksnumbered=true,    % Numbers in PDF bookmarks
            bookmarksopen=true,        % Don't show bookmaks when opened
            hyperindex=true,           % Numbers in index are links
            pdfpagelabels=true,        % Set PDF page labels
            pdfa]{hyperref}                         % Klickable Links and ToC entries (should be last package !!!)

\hypersetup{
    pdfauthor    = {Thomas Gnädig, Etienne Gramlich, Tim Hardenacke, Merle Wolff},
    pdftitle     = {DeepRain},
    pdfsubject   = {Deep Learning},
    pdfkeywords  = {Deep Learning,rain},
    pdflang      = {de},         % Set language of document to German
    pdfstartview = {FitH},       % Fit the document to screen horizontally (page width)
    pdfview      = {FitH},       % Fit the document to screen horizontally (page width)
    colorlinks   = {true},       % Use color for links
    linkcolor    = {black},      % Links in black (text color)
    urlcolor     = {black},      % URLs in black (text color)
    citecolor    = {black},      % Citations in black (text color)
    filecolor    = {black},      % File links in black (text color)
}

\title{DeepRain}
\author{Thomas Gnädig, Etienne Gramlich, Tim Hardenacke, Merle Wolf}

\begin{document}
\maketitle
\tableofcontents
\newpage

\begin{acronym}[DWD] % Hier steht die längste Abkürzung, damit die Breite stimmt
 \acro{DWD}{Deutscher Wetterdienst}
 \acro{MSE}{Mean Squared Error}
\end{acronym}

%Einleitung zu DeepLearning
\input{Abstract.tex}

\input{Daten.tex}

% Datenaufbereitung
\input{Datenaufbereitung.tex}

% Datenanalyse
\input{Datenanalyse.tex}
% Netzwerke
\input{Architekturen.tex}
\input{mse.tex}
\input{Klassifizierung.tex}

% Ziel und Test: MSE

% Fazit / Ausblick
\input{fazit.tex}

\newpage
\listoffigures
\listoftables

\bibliography{paper/bib/report.bib}
\bibliographystyle{alpha}

\end{document}